---
title: "Combined Leaf Datasets"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

cell_data <- tar_read(leaf_sobj_celldata)
```

## Plot all cells from all studies

Here I plot the integrated embedding of all cells from 8 individual datasets, 
colored by predicted cell type.

```{r}
umap_base <- ggplot(cell_data %>% arrange(DC3000.Up1), aes(x = UMAP_1, y = UMAP_2)) + theme_bw()
umap_base + geom_point(aes(color = predicted.id), size = 1, alpha = 0.1) + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 3))) +
  coord_fixed() +
  theme_void() +
  theme(axis.title.x = element_text(hjust = 0)) +
  theme(axis.title.y = element_text(hjust = 0, angle = 90))

ggsave("plots/combined_leaf_umap_celltype.pdf", family = "ArialMT")
ggsave("plots/combined_leaf_umap_celltype.jpg", family = "ArialMT", dpi = 600)
```

Here I break down the integrated embedding by individual study

```{r}
umap_base + geom_point(aes(color = predicted.id), alpha = 0.1, size = 1) + 
  facet_wrap("Sample_Name") +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 3)))
ggsave("plots/combined_leaf_umap_celltype_sample.pdf", family = "ArialMT")
```

## Plot pathogen response on integrated embedding

Here I plot the DC3000 response score (expression of up-regulated genes minus 
down-regulated genes) on the integrated UMAP.

```{r}
umap_base + geom_point(aes(color = DC3000.Up1 - DC3000.Down1), size = 1) + 
  scale_color_viridis_c(option = "plasma")

ggsave("plots/combined_leaf_umap_response.pdf", family = "ArialMT")
```

## Plot proportion of cells from each dataset over a specific area

```{r}
sample_names <- cell_data$Sample_Name %>% unique()

plots <- map(sample_names, function(sample_name) {
  plt <- ggplot(cell_data %>% arrange(UMAP_1), aes(x = UMAP_1, y = UMAP_2, group = -1)) +
    stat_summary_hex(aes(z = Sample_Name == sample_name, color = Sample_Name == sample_name), fun = mean, binwidth = 0.2) +
    scale_fill_viridis_c(option = "plasma", name = paste0("Proportion ", sample_name, " Cells")) +
    labs(x = "UMAP 1", y = "UMAP 2") +
    coord_fixed() +
    theme_void() +
    theme(axis.title.x = element_text(hjust = 0)) +
    theme(axis.title.y = element_text(hjust = 0, angle = 90))
  
  ggsave(paste0("plots/combined_leaf_umap_sampledensity_", sample_name, ".pdf"), plot = plt, family = "ArialMT")
  
  plt
})

plots
```

