---
title: "Protoplast Effect"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

sobj <- tar_read(archived_sobj)

bulk_de_table <- tar_read(bulk_de_table) %>% 
  mutate(
    ProtoDE = factor(case_when(
      p_adj_proto < 0.05 & logFC_proto < 0 ~ "Proto Down",
      p_adj_proto < 0.05 & logFC_proto >= 0 ~ "Proto Up",
      p_adj_proto >= 0.05 ~ "Proto Not DE"
    ), levels = c("Proto Down", "Proto Up", "Proto Not DE"))
  ) %>% 
  mutate(
    TreatDE = factor(case_when(
      p_adj_treat < 0.05 & logFC_treat < 0 ~ "DC3000 Down",
      p_adj_treat < 0.05 & logFC_treat >= 0 ~ "DC3000 Up",
      p_adj_treat >= 0.05 ~ "DC3000 Not DE",
    ), levels = c("DC3000 Down", "DC3000 Up", "DC3000 Not DE"))
  )

proto_down <- filter(bulk_de_table, de_type_proto == "Proto Down") %>% pull(Locus)
proto_up <- filter(bulk_de_table, de_type_proto == "Proto Up") %>% pull(Locus)

sobj <- AddModuleScore(object = sobj, name = "Proto.Down", features = list(proto_down), assay = "SCT")
sobj <- AddModuleScore(object = sobj, name = "Proto.Up", features = list(proto_up), assay = "SCT")

cell_data <- get_cell_data(sobj) %>% 
  mutate(ProtoScore = Proto.Up1 - Proto.Down1) %>% 
  mutate(
    ProtoScoreSquish = scales::squish(
      ProtoScore, 
      range = c(
        quantile(ProtoScore, 0.01),
        quantile(ProtoScore, 0.99)
      )
    )
  ) %>% 
  mutate(Cluster_Type = factor(Cluster_Type, levels = c(
    "Other Mesophyll",
    "Immunity Mesophyll",
    "Transition Mesophyll",
    "Susceptibility Mesophyll",
    "Bundle Sheath",
    "Companion",
    "Vasculature",
    "Guard"
  ))) %>% 
  mutate(Cluster = factor(Cluster, levels = c(
    paste0("M", 1:14),
    paste0(c("B","P","G","C"), 15:18)
  )))
```

## Plot Protoplast Module Score on UMAP

```{r}
color_scheme <- tar_read(archived_clusterNames) %>% 
  select(Cluster_Type, Color) %>% 
  deframe()

ggplot(arrange(cell_data, UMAP_1), aes(x = UMAP_1, y = UMAP_2)) + 
  geom_point(aes(color = ProtoScoreSquish), size = 0.5) + 
  scale_color_viridis_c(name = "Protoplast Signature Score", option = "plasma") +
  labs(x = "UMAP 1", y = "UMAP 2") +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank())

ggsave("plots/protoplast_module_umap.pdf", family = "ArialMT")
```

## Plot Protoplast Module Score by Cell Type (boxplot)

```{r}
color_scheme <- tar_read(archived_clusterNames) %>% 
  select(Cluster, Color) %>% 
  deframe()

ggplot(cell_data, aes(x = Cluster, y = ProtoScore)) + 
  geom_boxplot(aes(fill = Cluster), outlier.shape = NA, size = 0.5) +
  geom_jitter(alpha = 0.1) +
  scale_fill_manual(name = "Cell Type", values = color_scheme) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))  +
  labs(x = "Cell Type") +
  theme_bw() +
  theme(axis.text = element_text(size = 7))

ggsave("plots/protoplast_module_boxplot.pdf", family = "ArialMT")
```

## Table of Protoplast-affected loci vs. DC3000-affected loci

(DE = adjusted p-value < 0.05)
```{r}
bulk_de_table %>% 
  select(ProtoDE, TreatDE) %>% 
  table()
```

## UpSet of Protoplast-affected loci vs. DC3000-affected loci

```{r, results='hide'}
gene_list <- list(
  Proto_Down = bulk_de_table %>% filter(ProtoDE == "Proto Down") %>% pull(Locus),
  Proto_Up = bulk_de_table %>% filter(ProtoDE == "Proto Up") %>% pull(Locus),
  DC3000_Down = bulk_de_table %>% filter(TreatDE == "DC3000 Down") %>% pull(Locus),
  DC3000_Up = bulk_de_table %>% filter(TreatDE == "DC3000 Up") %>% pull(Locus)
)

UpSetR::fromList(gene_list) %>% UpSetR::upset()

pdf(file="plots/de_upset_dc3000_v_proto.pdf")
UpSetR::fromList(gene_list) %>% UpSetR::upset()
dev.off()
```

